### 2022-03-14 paper/output/document/thesis_5_discussion.tex
CLEANING DONE

%
\begin{isabellebody}%
\setisabellecontext{thesis{\isacharunderscore}{\isadigit{5}}{\isacharunderscore}discussion}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Discussion\label{discussion}%
}
\isamarkuptrue%
%
\isamarkupsubsection{Automated Moral Agents in Practice \label{AIethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In Chapter \ref{applications}, I demonstrated that my system is capable of performing sophisticated,
nuanced ethical reasoning. In this section, I outline the additional components necessary for my system
to guide AI agents in practice. As it stands, my system is a categorical imperative library that takes 
as input the logical representation of a maxim and returns its moral status (if it is obligatory, 
prohibited, or permissible). My project potentially 
serves as one component of an ``ethics engine'' that an AI agent could use to make ethical decisions.
For example, my system could be combined with an input parser to translate moral dilemmas as represented 
to the AI agent into maxims in my logic. An output parser could translate my system's output
into a prescription for action that the AI agent could act on. Figure \ref{fig:AIengine} depicts the 
workflow of this example ethics engine.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{center}
\begin{figure}
\includegraphics[scale=0.43]{AI_engine.png}
\caption{An example of an ethics engine for an artificial agent, which begins with a moral dilemma, 
passes it through an input parser, applies the automated categorical imperative test, and then processes the 
output using an output parser. I contribute the automated categorical imperative component.} \label{fig:AIengine}
\end{figure}
\end{center}
%
\begin{isamarkuptext}%
In this workflow, an AI agent is faced with a moral dilemma in some internal representation. The input
parser translates this internal representation into an appropriate logical representation, i.e. 
a circumstance, act, goal tuple. The output parser translates the output of the categorical imperative
library (the moral status of the maxim as obligatory, prohibited, or permissible) to a prescription for
action. In order for my system to be used in an AI agent using the workflow
above, future work would need to develop such input and output parsers.

The input parser must translate a complex real-world situation into a flat, logical representation.
This requires that the input parser determine which circumstances are morally relevant
to a maxim, a controversial judgement. As mentioned in Section \ref{whatisamaxim},
there is robust debate in the literature on the circumstances that should be considered when formulating a maxim, 
inspired by the tailoring objection, a common criticism of Kantian ethics. Recall that the tailoring objection is the worry that arbitrarily specific 
circumstances render any maxim universalizable. For example, consider the maxim ``When my name is Lavanya Singh 
and I am wearing a purple shirt and it is November 26th, I will lie in order to get some easy cash.'' 
Even if this maxim is willed universally, the circumstances are so 
specific that lying will not become the general mechanism for getting easy cash, so the lender will 
believe my lie and the maxim will remain effective. By tailoring the circumstances, any maxim can 
evade universalization.

The Kantian response to this criticism is to require that the circumstances included in the formulation
of the maxim be morally relevant. In the example above, my purple shirt and the date clearly have no bearing on 
the moral status of lying. On the other hand, consider the maxim, ``When I am unemployed, I will murder
someone in order to take their job.'' The circumstances of being unemployed clearly have some bearing on the moral
relevance of the murder in question; they speak to the motivation for the murder.

While this view has intuitive appeal, it raises the question of how we can determine
which circumstances are morally relevant. O'Neill answers this question by noting that the Formula of Universal Law is 
a ``test of moral worth rather than of outward rightness'' \citep[98]{constofreason}. The FUL is a way 
for an agent to decide how they should behave, not for a third-party to judge their behavior. Ethics is 
a personal process and the FUL is designed to help agents make decisions. Because agents use 
the FUL to evaluate their own behavior, the test is at its 
best when they make a good faith effort to isolate the \emph{principle} of their action, rather than some
``surface intent'' \citep[87]{constofreason}. The FUL is supposed to determine if an agent's principle of action
is universally consistent, so it is most effective when an agent accurately formulates the principle
they act on. Circumstances are morally relevant if they reflect the way that the agent is 
thinking about their own action. In the example above, the circumstance of wearing a purple shirt doesn't reflect
the principle of the liar's action. Its inclusion is a disingenous attempt to evade the universalizability
test, but because the FUL is a test of personal integrity, it cannot withstand this kind of mental
gymnastics.

While the above account how a well-intentioned human agent can determine 
morally relevant circumstances, the challenge remains open for automated ethics. In order for an automated
ethical agent to use the categorical imperative to its fullest extent, the input maxim must be a good-faith 
attempt to capture the agent's principle
of action. However an action is turned into a maxim for my system to process, whether manually as I did
in Chapter \ref{applications} or using an automatic input parser, this transformation must be a good-faith 
attempt to capture the principle of action. Translating everyday situations into appropriate maxims is 
the bulk of the work that a Kantian human 
being does when making decisions. Common misconceptions about Kantian ethics\footnote{For example, critics 
of Kantian ethics worry that the maxim, ``When I am a
man, I will marry a man because I want to spend my life with him'' fails the universalizability
test because if all men only married men, sexual reproduction would stop. This argument implies 
that Kantian ethics is homophobic. Kantians often respond by arguing that the correct formulation of 
this maxim is, ``When I love a man, I will marry him because I want to spend my life with him,'' which
is universalizable because if everyone marries who they love, some men will marry women and others will
marry men.} often result from incorrectly formulated maxims, 
and the entire field of applied Kantian ethics is devoted to generating the right kinds of maxims to test.

The challenge of formulating a maxim is one of the biggest obstacles to using my categorical imperative library
in an AI ethics engine. One solution is for a human being to perform the role of the input
parser. Once an AI agent stumbles onto an ethical dilemma, a human being could take over, formulate 
the right question, and feed it into the categorical imperative library to see what action the categorical 
imperative would prescribe. For proponents of the ``human-in-the-loop'' model of AI ethics, in which 
ethical AI requires that humans guide machines, this kind of human involvement may be a feature \citep{loop}.
The outcome of the universalizability test will depend on how the human formulates the maxim; if the 
human puts garbage into the test, the test will return garbage out.

It is likely that, regardless of the strengths of the human-in-the-loop model, fully automated AI 
agents will exist. Even if developing this kind of AI is irresponsible,
such developments are likely and will require ethics engines, or risk no consideration of ethics at all.
In such a world, the input parser in my ethics engine would have to be automated.
It is likely that, just as implementations of automated ethics choose 
a particular ethical theory and implement it, different implementations of such an input parser would 
adopt different interpretations of maxim formulation and morally relevant circumstances.

These interpretations could inspire heuristics to classify circumstances as morally 
relevant. For example, one such attempt could define a moral closeness relation between an action, a 
goal, and circumstances. This heuristic could define morally relevant circumstances as those that 
reach a certain closeness threshhold with the action and the goal. Another possible heuristic could 
define some set of morally important entities, and classify morally relevant circumstances as those
that involve morally important entities. I discuss a potential machine-learning based approach which formulates
maxims based on a training set of appropriately formulated maxims in Section \ref{amapossible}.
Determining morally relevant circumstances, either using heuristics or human involvement, is a ripe 
area for future work.

Once the input has been parsed, either by a human or a machine, into  a sentence in my logic, my 
project can evaluate its moral status using my implementation of 
the FUL. Concretely, my project returns a value indicating if the maxim is obligatory, permissible, 
or prohibited. The maxim is prohibited if it fails the universalizability test, permissible if it passes, and obligatory 
if its negation fails the universalizability test. All three of these properties require testing if a 
certain theorem holds or not in my logic, a calculation that I demonstrate in Section \ref{testing}. 
Testing these properties requires that my system have a database of common sense or factual background. 
Different applications of my system may require different factual background (e.g. a self-driving car 
would need to know traffic regulations), so this common sense database may need to be application 
specific. As demonstrated in the examples in Chapter \ref{applications}, my system can produce sophisticated 
judgements with relatively little situational context. While the need for factual background is a challenge
for automated ethics, Chapter \ref{applications} demonstrates that it is less daunting than it seems.

@@

The ethics engine depicted above is a high-level example of one way to use my project to guide an artifical agent,
with additional work to parse the input and output of my implementation of the categorical imperative.
Effectively, the kind of automated ethics I implement could be part of a library that AI developers use to 
give AI agents the capacity for sophisticated ethical reasoning faithful to philosophical literature. 
This represents an improvement over existing AI ethics, which rarely captures the complexity 
of any ethical theory that philosophers plausibly defend.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Computational Ethics \label{computationalethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In addition to guiding AI agents, automated ethics can also help human beings make philosophical
progress. Just as theorem provers make mathematics more efficient and push mathematicians to think 
precisely about the phenomena they model, computational ethics can help philosophers ask and answer
new philosophical questions. In this section, I share an example of the kind of philosophical insight that computational ethics 
can prompt and analyze the value that this tool can offer to philosophers.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Example of a Philosophical Insight: Well-Formed Maxims%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
As presented in Section \ref{formalizingful}, in the process of developing my formalization of
the FUL, I discovered that certain kinds of maxims are badly formed, or inappropriate inputs to the 
universalizability test. The FUL is consistent only if it holds for ``well-formed maxims,''
such that neither the act nor goal are already achieved in the given circumstances. Precisely, 
a circumstance, act, goal tuple (c, a, g) is well-formed if $(\neg (c \longrightarrow a) ) \wedge 
(\neg(c \longrightarrow g))$. The insight that the FUL should not apply to badly-formed maxims has 
philosophical value and serves as evidence of the potential of computational ethics.

\noindent \textbf{Philosophical Implications of Badly-Formed Maxims}

Isabelle gave a logical argument for why the FUL can only hold for well-formed maxims, and I return to Kantian
literature to better understand the philosophical implications of this idea. Because badly-formed maxims
neither change an agent's behavior nor generate meaningful obligations, they are not the right kinds of 
actions for practical reasoners to make moral judgements about. They cannot be action-guiding and are thus not the kind of problem that 
ethics should be concerned with. Moreover, under the Kantian account of the will, the very act of asking 
if a badly-formed maxim is prohibited generates a contradiction by undermining the will's authority over itself.

Consider the example badly-formed maxim, ``When eating breakfast, I will eat breakfast in order to 
eat breakfast.'' There is something empty about this maxim because acting on it could never result in 
any action. If I adopts this maxim as a principle to live by,
I decide that, in the circumstances ``eating breakfast'' I will perform the act ``eating breakfast''
for the purpose ``eating breakfast.'' In these circumstances, the act has 
already been performed. Adopting this maxim as a law to live by does not change how I live. If I adopt 
this maxim, then when I am eating breakfast, I eat breakfast, but this statement is already tautologically true.

@@

Because badly-formed maxims can't prescribe or alter action, they are not practically action-guiding and 
thus are not the right kinds of maxims for practical reasoners to evaluate. Insofar as ethics 
is supposed to guide action, badly-formed maxims cannot be part of this project because they
have no bearing on what someone should do. Practical reason is the kind of reason that helps us decide 
what we should do. A practical reasoner asks moral questions not as a mental puzzle or out of curiosity, but 
to decide how to act. A badly-formed maxim is not the kind of maxim that a practical reasoner should consider, because it
will have no bearing on what the agent should do. It makes no sense of ask, ``Should I act on a badly-formed
maxim?'' because the answer to this question cannot change your behavior. There is no explicit prohibition 
against a badly-formed maxim, but it is the wrong kind of question for a practical reasoner to ask.

@@

Above, I defined badly formed maxims, a philosophical concept that I discovered
using computational ethics. In this section, I explore the implications of this concept for a 
philosophical debate to show that computational ethics can result in insights with real 
philosophical weight. I use badly formed maxims to resolve the tension between self-doubt and self-respect. 
This debate originated in epistemology, which values doubting your beliefs but also rationally requires
that you believe that you are not mistaken (else you should update your beliefs). 
Self-doubt and self-respect also compete in ethics, where questioning your judgements is valuable but 
questioning a commitment as you make it is impossible. Self-doubt and self-respect appear 
irresolvably opposed until they are understood through the lens of badly formed maxims. I argue that naive
conceptions of self-doubt are badly formed maxims in disguise. If we reformulate these maxims to be well-formed,
the tension between self-doubt and self-respect dissolves, and it becomes clear that both concepts can
co-exist. I sketch the details of this argument in the rest of this section. I will conclude that 
badly-formed maxims serve as one guide to the formulation of maxims, and thus may have bearing on 
other debates in Kantian ethics as well. I first introduce the tension between self-doubt and self-respect 
in epistemology, then explain the parallel tension in ethics, and finally present a resolution of this 
tension using badly-formed maxims.

In epistemology, there is a tension between the rational requirement to believe in yourself and the 
value of self-doubt, in moderation. Christensen presents the ``principle of self-respect,'' which requires 
that a rational agent refrain from believing that they have mistaken beliefs \cite[4]{christensen}. For example, I cannot 
rationally believe both that the sky is blue and that I believe that the sky is green. In other words, I cannot 
disapprove of my own credences, since if I do disapprove of them, I would just abandon them. Christensen 
argues that this principle, which he abbreviates to SR, holds because 
a perfectly rational agent can make accurate and confident judgements about what they believe. If this 
is the case, violating SR results in a simple contradiction \cite[8-9]{christensen}.

While most philosophers accept some version of SR\footnote{Christensen cites \citet{vanfraassen}, 
\citet{vickers}, and \citet{koons}.}, 
Roush argues that the principle must be modified in order to account for healthy epistemic 
self-doubt. She argues that, while pathological second-guessing is correctly criticized, we are generally 
imperfect beings, and some sensitivity to our own limitations is a virtue \cite[2]{roushselfhelp}. Even Christensen 
acknowledges that total self-confidence is an epistemic flaw \cite[1]{christensen}. Thus, there is tension between the rational
requirement to respect our authority as believers and the practical reality that we are often wrong.

@@

To resolve this tension, I return to my earlier example of a dancer. Imagine Sara is dancing at a 
weddding, when, in a moment of angst, she asks herself, ``Should I really be dancing right now?'' 
It seems that she asking if the maxim, ``When dancing at your friend's wedding, dance for the sake 
of dancing'' is a permissible maxim to act on. Notice that the maxim in question is badly-formed: the 
circumstance ``when dancing at a friend's wedding'' implies the act ``dance.'' As I argued above, if 
Sara asks if she should will a badly formed maxim, she is questioning her own will's authority, which
is paradoxical. If maxims are about self-doubt are badly-formed maxims, then the tension between ESR
and ESD is natural and unavoidable: debating the permissibility of a badly-formed maxim inherently
involves questioning commitments as we make them, which is impossible. This is the source of the 
tension between ESR and ESD. Those committed to this 
interpretation must abandon one principle or the other, since committing and questioning are incompatible.

If ethical self-doubt as badly-formed maxim contradicts self-respect, then resolving this issue requires a 
different interpretation of ethical self-doubt. Under this interpretation, 
when Sara asks, ``Should I really be dancing right now?'' she wants to know if the maxim that 
resulted in the current moment of dancing was the right thing to will. She is 
asking if she made the right decision in the past, when she decided to dance. The maxim that initiated 
the dancing may be something like ``When at a wedding, dance for the sake of dancing.'' This is the maxim 
that she is currently acting on, not the badly-formed maxim  from above. Under this interpretation, 
there is no tension at all between self-doubt and self-respect. It is perfectly valid for a reasoner 
to doubt their prior moral judgements, just as it is perfectly rational for a believer to doubt their 
past beliefs \cite[3-4]{christensen}. Such doubt does not undermine the reasoner's decision-making 
capacity and is thus perfectly consistent with ethical self-respect. Moreover, this is a much more 
intuitive question to be asking \emph{because} it is a well-formed maxim. Not only is the badly-formed
interpretation of this maxim problematic for ESR, it is also not the kind of question that is useful for a practical
reasoner to ask, as argued above.

@@

Computational ethics is at its infancy. The use of theorem provers in mathematics is just now beginning 
to make headway \citep{buzzardvideo}, even though theorem provers were first invented in the 1960's \citep{historyofITP}. 
In contrast, the first attempts to use theorem provers for progress in any kind of philosophy occurred 
in the last few years \color{red} and, at the time of this writing, no known work has used theorem 
provers for computational ethics \color{black}. The 
fact that this nascent technology is already helping humans reach non-trivial philosophical conclusions 
is reason to, at the very least, entertain the possibility of a future where computational ethics 
becomes as normal for philosophers as using a thought experiment.

>>>>

%
\begin{isabellebody}%
\setisabellecontext{thesis{\isacharunderscore}{\isadigit{5}}{\isacharunderscore}discussion}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Discussion\label{discussion}%
}
\isamarkuptrue%
%
\isamarkupsubsection{Automated Moral Agents in Practice \label{AIethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{center}
\begin{figure}
\includegraphics[scale=0.43]{AI_engine.png}
\caption{An example of an ethics engine for an artificial agent. This ethics engine passes a moral dilemma 
through an input parser, applies the automated categorical imperative test, and finally processes the 
output using an output parser, producing a prescription for action. I contribute the automated categorical 
imperative component.} \label{fig:AIengine}
\end{figure}
\end{center}
%
\begin{isamarkuptext}%
In Chapter \ref{applications}, I demonstrated that my system is capable of performing sophisticated,
nuanced ethical reasoning. In this section, I outline the additional components necessary for my system
to guide AI agents in practice. As it stands, my system is a categorical imperative library that, given
sufficient factual or ``common sense'' background, can take 
as input the logical representation of a maxim and return its moral status (if it is obligatory, 
prohibited, or permissible). My project potentially 
serves as one component of an ``ethics engine'' that an AI agent could use to make ethical decisions.
For example, my system could be combined with an input parser to translate moral dilemmas as represented 
to the AI agent into maxims in my logic. An output parser could translate my system's output
into a prescription for action that the AI agent could act on. Figure \ref{fig:AIengine} depicts the 
workflow of this example ethics engine.

In this workflow, an AI agent is faced with a moral dilemma in some internal representation. The input
parser translates this internal representation into an appropriate logical representation, i.e. 
a circumstance, act, goal tuple. The output parser translates the output of the categorical imperative
library (the moral status of the maxim as obligatory, prohibited, or permissible) to a prescription for
action. In order for my system to be used in an AI agent using the workflow
above, future work must develop such input and output parsers.

One of the biggest challenges for an ethics engine is the development of an input parser. An input parser 
for my implementation of automated Kantian ethics must translate a complex real-world situation into a flat, logical representation.
This requires that the input parser determine which circumstances are morally relevant
to a maxim, a controversial judgement. As mentioned in Section \ref{whatisamaxim},
there is robust debate on the circumstances that should be considered when formulating a maxim, 
inspired by a common criticism of Kantian ethics called the tailoring objection. Recall that the 
tailoring objection is the worry that arbitrarily specific 
circumstances render any maxim universalizable. For example, consider the maxim ``When my name is Lavanya Singh 
and I am wearing a purple shirt and it is November 26th, I will lie in order to get some easy cash.'' 
Even if this maxim is willed universally, the circumstances are so 
specific that lying will not become the general mechanism for getting easy cash, so the lender will 
believe my lie and the maxim will remain effective. By tailoring the circumstances, any maxim can 
evade universalization.

The Kantian response to this criticism is to require that the circumstances included in the formulation
of the maxim be morally relevant. In the example above, my purple shirt and the date have no bearing on 
the moral status of lying. On the other hand, consider the maxim, ``When I am unemployed, I will murder
someone in order to take their job.'' The circumstances of being unemployed clearly have some bearing on the moral
relevance of the murder in question; they speak to the motivation for the murder.

While this view has intuitive appeal, it raises the question of how we can determine
which circumstances are morally relevant. O'Neill answers this question by noting that the Formula of Universal Law is 
a ``test of moral worth rather than of outward rightness'' \citep[98]{constofreason}. The FUL is a way 
for an agent to decide how they should behave, not for a third-party to judge their behavior. Ethics is 
a personal process and the FUL is designed to help agents make decisions for themselves. Because agents use 
the FUL to evaluate their own behavior, the test is at its 
best when they make a good faith effort to isolate the \emph{principle} of their action, rather than some
``surface intent'' \citep[87]{constofreason}. The FUL is supposed to determine if an agent's principle of action
is universally consistent, so it is most effective when an agent accurately formulates the principle that
they act on. Circumstances are morally relevant if they reflect the way that the agent is 
thinking about their own action. In the example above, the circumstance of wearing a purple shirt doesn't reflect
the principle of the liar's action. Its inclusion is a disingenous attempt to evade the universalizability
test, but because the FUL is a test of personal integrity, it cannot withstand this kind of mental
gymnastics.

While the above account explains how a well-intentioned human agent can determine 
morally relevant circumstances, the challenge remains open for automated ethics. However an action is turned into a maxim for my system 
to process, whether manually as I did in Chapter \ref{applications} or using an automatic input 
parser, this transformation must be a good-faith attempt to capture the principle of action. 
Correctly formulating a maxim requires what Kantians call ``practical judgement,'' or
common sense reasoning and factual background \citep{oneilluniversallaws}. Returning to the example above, noticing that being 
unemployed may contribute to one's desire to steal another's job requires practical
judgement, not just pure reason alone. Translating everyday situations into appropriate maxims is
the bulk of the reasoning that a Kantian human being does when making decisions. Automating this
reasoning requires endowing a machine with common sense.

The challenge of formulating a maxim is one of the biggest obstacles to using my categorical imperative library
in an AI ethics engine. One solution is for a human being to perform the role of the input
parser by supervising the operation of an AI agent. When the agent stumbles onto
an ethical dilemma, the human could take over, formulate
the right question, and feed it into the categorical imperative library to see what action the categorical 
imperative would prescribe. Alternatively, AI developers can identify expected ethical dilemmas 
that the machine may face in advance and hardcode their own judgements for these dilemmas. 
For proponents of the ``human-in-the-loop'' model of AI ethics, in which
ethical AI requires that humans guide machines, this kind of human involvement may be a feature \citep{loop}.
Both of these solutions imply that the outcome of the universalizability test will depend on how the 
human formulates the maxim; if the human puts garbage into the test, the test will return garbage out.

It is likely that, regardless of the strengths of the human-in-the-loop model, fully automated AI 
agents will exist. Even if developing this kind of AI is irresponsible,
such developments are likely and will require ethics engines, or risk no consideration of ethics at all.
In such a world, the input parser in my ethics engine would have to be automated.
It is likely that, just as implementations of automated ethics choose 
a particular ethical theory and implement it, different implementations of such an input parser may 
adopt different interpretations of maxim formulation and morally relevant circumstances.

These interpretations could inspire heuristics to classify circumstances as morally 
relevant. For example, one such attempt could define a moral closeness relation between an action, a 
goal, and circumstances. This heuristic could define morally relevant circumstances as those that 
reach a certain closeness threshhold with the action and the goal. Another possible heuristic could 
define some set of morally important entities, and classify morally relevant circumstances as those
that involve morally important entities. I discuss a potential machine-learning based approach which formulates
maxims based on a training set of appropriately formulated maxims in Section \ref{amapossible}. This 
approach mimics how human beings formulate maxims; we use common sense and prior situational, 
factual, and ethical knowledge to isolate our principle of action. Determining morally relevant circumstances, 
either using heuristics or human involvement, is a ripe area for future work.

Once the input has been parsed, either by a human or a machine, into a sentence in my logic, my 
project can evaluate its moral status using my implementation of 
the FUL. Concretely, my project returns a value indicating if the maxim is obligatory, permissible, 
or prohibited. The maxim is prohibited if it fails the universalizability test, permissible if it passes, and obligatory 
if its negation fails the universalizability test. All three of these properties require testing if a 
certain theorem holds or not in my logic, a calculation that I demonstrate in Section \ref{testing}. 
Testing these properties requires that my system have a database of common sense or factual background. 
Different applications of my system may require different factual background (e.g. a self-driving car 
needs to know traffic regulations), so this common sense database will need to be application 
specific. As demonstrated in the examples in Chapter \ref{applications}, my system can produce sophisticated 
judgements with relatively little situational context. Thus, while my automated categorical imperative
requires some common sense and factual background, Chapter \ref{applications} demonstrates that automating
this common sense is less daunting than it seems.

@@

The ethics engine depicted above is a high-level example of one way to use my project to guide an artifical agent,
with additional work to parse the input and output of my implementation of the categorical imperative.
Effectively, the kind of automated ethics I implement could be part of a library that AI developers use to 
give AI agents the capacity for sophisticated ethical reasoning faithful to philosophical literature. 
This represents an improvement over existing automated ethics, which rarely captures the complexity 
of any ethical theory that philosophers plausibly defend.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Computational Ethics \label{computationalethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In addition to guiding AI agents, automated ethics can also help human beings make philosophical
progress. Just as theorem provers make mathematics more efficient and push mathematicians to think 
precisely about the phenomena they model, computational ethics can help philosophers ask and answer
new philosophical questions. In Section \ref{joking}, I presented one example of the power of computational ethics
by using my system to isolate a necessary and sufficient condition for the act of lying, or knowingly saying
something false, to be wrong. In this section, I share another example of the kind of philosophical 
insight that computational ethics can prompt and analyze the value that this tool can offer to philosophers.\footnote{
I present a final example of computational ethics in Appendix \ref{weirdtests}, where I resolve an ambiguity 
in Korsgaard's argument for the wrongness of false promising using my system \citep{KorsgaardFUL}.} 
Fields from protein folding to game theory are uncovering new insights using computational tools, and computational
ethics harnesses this power to inspire similar progress in philosophy.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Example of a Philosophical Insight: Well-Formed Maxims%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
As presented in Section \ref{formalizingful}, in the process of developing my formalization of
the FUL, I discovered that certain kinds of maxims are badly formed, or inappropriate inputs to the 
universalizability test. The FUL is consistent only if it holds for ``well-formed maxims,''
such that neither the act nor goal are already achieved in the given circumstances. Precisely, 
a circumstance, act, goal tuple (c, a, g) is well-formed if $(\neg (c \longrightarrow a) ) \wedge 
(\neg(c \longrightarrow g))$, or if the circumstances do not imply the act or the goal. The insight
that the FUL does not apply to badly-formed maxims has philosophical value and serves as evidence 
of the potential of computational ethics. In the next section, I explain and explore this insight using 
Kantian literature. In the following section, I demonstrate the philosophical implications of well-formed
maxims by using this insight to resolve a tension between self-doubt and self-respect. The idea of well-formed
maxims can guide the formulation of maxims and thus may have implications for many parts of Kantian ethics.

\noindent \textbf{Badly-Formed Maxims and Kantian Ethics}

Isabelle gave a logical argument for why the FUL can only hold for well-formed maxims, and I return to Kantian
literature to better understand this idea. In this section, I argue that 
because badly-formed maxims
neither change an agent's behavior nor generate meaningful obligations, they are not the right kinds of 
actions for practical reasoners to make moral judgements about. They cannot be action-guiding and are thus not the kind of problem that 
ethics should be concerned with. Moreover, under the Kantian account of the will, the very act of asking 
if a badly-formed maxim is prohibited generates a contradiction by undermining the will's authority over itself.

Consider the example badly-formed maxim, ``When eating breakfast, I will eat breakfast in order to 
eat breakfast.'' There is something empty about this maxim because acting on it could never result in 
any action. If I adopt this maxim as a principle to live by,
I decide that, in the circumstances ``eating breakfast'' I will perform the act ``eating breakfast''
for the purpose ``eating breakfast.'' In these circumstances, the act has 
already been performed. Adopting this maxim as a law to live by does not change how I live. If I adopt 
this maxim, then when I am eating breakfast, I eat breakfast, but this statement is already tautologically true.

@@

Because badly-formed maxims can't prescribe or alter action, they are not practically action-guiding and 
thus are not the right kinds of maxims for practical reasoners to evaluate. Insofar as ethics 
is supposed to guide action, badly-formed maxims cannot be part of this project because they
have no bearing on what someone should do. Practical reason is the kind of reason that helps us decide 
what we should do. A practical reasoner asks moral questions not as a mental puzzle or out of curiosity, but 
to decide how to act. A badly-formed maxim is not the kind of maxim that a practical reasoner should consider, because it
will have no bearing on what the agent should do.

@@

Above, I defined badly formed maxims, a philosophical concept that I discovered
using computational ethics. In this section, I explore the implications of this concept for the ethical
tension between self-doubt and self-respect to show that computational ethics can result in insights with real 
philosophical weight. The debate between self-doubt and self-respect originates in epistemology, which 
values doubting your beliefs but also rationally requires that you believe that you are not mistaken 
(else you should update your beliefs). I present a parallel tension between self-doubt and self-respect
in ethics, where questioning your judgements is valuable but 
questioning a commitment as you make it is impossible. Ethical self-doubt and self-respect appear 
irresolvably opposed until they are understood through the lens of badly formed maxims. I argue that naive
conceptions of self-doubt are badly formed maxims in disguise. If we reformulate these maxims to be well-formed,
the tension between self-doubt and self-respect dissolves. I sketch the details of this argument in 
the rest of this section. I first introduce the tension between self-doubt and self-respect 
in epistemology, then explain the parallel tension in ethics, and finally present a resolution of this 
tension using badly-formed maxims. I conclude that well-formed maxims form a boundary condition for
the formulation of maxims, so they may resolve many debates in Kantian literature and ethics more generally.

In epistemology, there is a tension between the rational requirement to believe in yourself and the 
value of self-doubt, in moderation. Christensen presents the ``principle of self-respect,'' which requires 
that a rational agent refrain from believing that they have mistaken beliefs \cite[4]{christensen}. For example, I cannot 
rationally believe both that the sky is blue and that I believe that the sky is green. In other words, I cannot 
disapprove of my own credences, since if I do disapprove of them, I should just abandon them. Christensen 
argues that this principle, which he abbreviates to SR, holds because 
a perfectly rational agent can make accurate and confident judgements about what they believe. If this 
is the case, violating SR results in a simple contradiction \cite[8-9]{christensen}.

While most philosophers accept some version of SR,\footnote{Christensen cites \citet{vanfraassen}, 
\citet{vickers}, and \citet{koons}.}
Roush argues that the principle must be modified in order to account for healthy epistemic 
self-doubt. She argues that, while pathological second-guessing is correctly criticized, we are generally 
imperfect beings, and some sensitivity to our own limitations is a virtue \cite[2]{roushselfhelp}. Even Christensen 
acknowledges that total self-confidence is an epistemic flaw \cite[1]{christensen}. Thus, there is tension between the rational
requirement to respect our authority as believers and the practical reality that we are often wrong.

@@

I now argue that well-formed maxims can resolve this tension. As in my earlier example, imagine that Sara is dancing at a 
weddding, when, in a moment of angst, she asks herself, ``Should I really be dancing right now?'' 
It seems that she asking if the maxim, ``When dancing at your friend's wedding, dance for the sake 
of dancing'' is a permissible maxim to act on. This maxim is badly-formed: the 
circumstance ``when dancing at a friend's wedding'' implies the act ``dance.'' As I argued above, if 
Sara asks if she should will a badly formed maxim, she is questioning her own will's authority, which
is paradoxical. If expressions of self-doubt are badly-formed maxims, then the tension between ESR
and ESD is natural and unavoidable: debating the permissibility of a badly-formed maxim inherently
involves questioning commitments as we make them, which is impossible. This is the source of the 
tension between ESR and ESD. Those committed to this 
interpretation must abandon one principle or the other, since committing and questioning are incompatible.

Because understanding ethical self-doubt as a badly-formed maxim contradicts self-respect, resolving this issue requires a 
different interpretation of ethical self-doubt. Under this interpretation, 
when Sara asks, ``Should I really be dancing right now?'' she wants to know if the maxim that 
resulted in the current moment of dancing was the right thing to will. She is 
asking if she made the right decision in the past, when she decided to dance. The maxim that initiated 
the dancing may be something like ``When at a wedding, dance for the sake of dancing.'' This is the maxim 
that she is currently acting on, not the badly-formed maxim  from above. Under this interpretation, 
there is no tension at all between self-doubt and self-respect. It is perfectly valid for a reasoner 
to doubt their prior moral judgements, just as it is perfectly rational for a believer to doubt their 
past beliefs \cite[3-4]{christensen}. Such doubt does not undermine the reasoner's decision-making 
capacity and is thus perfectly consistent with ethical self-respect. Moreover, this is a much more 
intuitive question to be asking \emph{because} it is a well-formed maxim. Not only is the badly-formed
interpretation of this maxim problematic for ESR, it is also not the kind of question that is useful for a practical
reasoner to ask, as argued above.

@@

Computational ethics is at its infancy. The use of theorem provers in mathematics is only now beginning 
to make headway \citep{buzzardvideo}, even though theorem provers were first invented in the 1960's \citep{historyofITP}. 
In contrast, the first attempts to use theorem provers to automate deontic logic occurred in the last
few years. The 
fact that this nascent technology is already helping humans reach non-trivial ethical conclusions 
is reason to, at the very least, entertain the possibility of a future where computational ethics 
becomes as normal for philosophers as using a thought experiment.
